---
title: "mlbgameday: Search Games"
author: "Kris Eberwein"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mlbgameday: Search Games}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include=TRUE}
```

In many cases, we may want to search for a specific player, stadium, or team. The `search_gids()` function provide limited functionality for such searches. Arguments include; team=NULL, start=NULL, end=NULL, venue=NULL, game_type=NULL, home_only=FALSE, away_only=FALSE.

# Examples

Find only regular season games for the 2016 season.

```{r, eval=FALSE}
library(mlbgameday)

gids <- search_gids(start = "2016-01-01", end = "2017-01-01", game_type = "r")

```

Find only regular season games played at Wrigley Field in 2016.

```{r, eval=FALSE}
library(mlbgameday)

gids <- search_gids(start = "2016-01-01", end = "2017-01-01", game_type = "r", venue = "Wrigley Field")

# This also works with a "fuzzy search"

gids <- search_gids(start = "2016-01-01", end = "2017-01-01", game_type = "r", venue = "Wrigley")


```

## Game Types

The `game_type = ` argument accects the following:

* "r" = regular season

* "d" = divisional playoffs

* "l" = leauge playoffs

* "w" = World Series

* "a" = all-star games

* "s" = spring training

* "e" = exabition game

Find only home games played by the Cleveland Indians in the regular season (regardless of stadium).

```{r, eval=FALSE}
library(mlbgameday)

gids <- search_gids(team = "indians", start = "2015-01-01", end = "2015-01-01", game_type = "r")

```

## Venue

The `search_gids()` function performs a "fuzzy match," on names of teams and ball parks. For example, we don't have to type out "Oriole Park at Camden Yards," simply "Camden" or "Camden Yards" will work.

```{r, eval=FALSE}
library(mlbgameday)

# This workds.

gids <- search_gids(team = "orioles", start = "2016-01-01", end = "2017-01-01", venue = "Oriole Park at Camden Yards")

# This also works.

gids <- search_gids(team = "orioles", start = "2016-01-01", end = "2017-01-01", venue = "Camden")


```

Be careful about venue changes or venue name changes. For example, "Marlins Park" opened in 2012, so the follwoing will not work.

```{r, eval=FALSE}
library(mlbgameday)

gids <- search_gids(team = "marlins", start = "2010-01-01", end = "2011-01-01", venue = "Marlins Park")


```

Instead, we would have to use the name of the old stadium. If we're not sure, the safest way would be to use `home_only = TRUE`.

```{r, eval=FALSE}
library(mlbgameday)

gids <- search_gids(team = "marlins", start = "2010-01-01", end = "2011-01-01", venue = "Sun Life Stadium")

# This will return the same result if we don't know the name of the stadium.

gids <- search_gids(team = "marlins", start = "2010-01-01", end = "2011-01-01", home_only = TRUE, game_type = "r")


```